Impact of grids and dynamical cores in CESM2.2 on the surface mass balance of the Greenland Ice Sheet

### 2022-09-10 second_submit/arcticgrids.tex
updates

{\color{blue}{Another potential factor in explaining the differences across the various grids and dycores is the surface albedo. Bare ice has a lower albedo than snow, so it's reasonable to speculate whether the large melt biases in the coarse grid simulations are in part due to more frequent bare ice exposure. One logical pathway to more bare ice exposure is related to the precipitation and cloud biases. As the coarser models accumulate too much precipitation in the GrIS interior, the coastal ice margins receive too little snow, which can be more easily melted away to expose the underlying glacial ice. Figure~\ref{fig:alb} shows maps of summer albedo biases in the simulations, relative to the RACMO2.3p2 and CERES-EBAF datasets. The surface albedo is diagnosed from the incident and net shortwave radiative fluxes at the surface. The lat-lon and quasi-uniform grids all have anomalously low albedo around the coasts in the southeast and northwest regions of the GrIS, which coincides with regions of anomalously low cloud cover and precipitation. In the VR simulations, the low albedo biases are alleviated to a large extent.

In light of these albedo results, one may interpret Figure~\ref{fig:bias} differently. That there are larger melt biases for grid cells with only only partial ice cover, may have less to do with the partial ice cover and more to do with the enhanced bare ice exposure in the coarser grids. Further experimentation is required to tease out the relative contributions of these two processes.}}

\begin{figure}[t]
\begin{center}
\begin{tabular}{cccc}
         \includegraphics[width=130mm]{temp_contours_diffALB.pdf}
\end{tabular}
\end{center}
\caption{1979-1998 northern hemisphere summer albedo bias (fraction), computed as the difference (top) from RACMO2.3p2, and (bottom) CERES-EBAF dataset. RACMO2.3p2 differences are found by mapping the model output to the individual model grids, and differences in the bottom panel are computed after mapping the individual model grids to the CERES-EBAF $1^{\circ}$ grid. Note that the averaging period for CERES-EBAF panels are 2003-2020, which is different from the averaging period for the model results.}
\label{fig:alb}
\end{figure}

@@

The $\tt Arctic-GrIS$ grid has the warmest summer lower troposphere of all grids, yet it has less melting than the quasi-uniform resolution SE grids. This suggests that grid resolution is somehow contributing to the melt biases in coarse grids, in a way that is not obvious from the large-scale dynamics. We propose a mechanism: coarse grids represent ablation zones using grid cells with mixed surface types, ice-covered and ice-free. The warmer ice-free patches may largely determine the mean state, leading to a warm bias over the ice-covered patches of the grid cell. This mechanism is supported by analysis of melt biases binned by grid-cell ice fraction.

>>>>

{\color{blue}{Another potential factor in explaining the melt differences across the grids and dycores is the surface albedo. Bare ice has a lower albedo than snow, so it's reasonable to speculate whether the large melt biases in the coarse grid simulations are in part due to more frequent bare ice exposure. One reasonable pathway to more bare ice exposure is related to the precipitation and cloud biases. As the lat-lon and quasi-uniform grids accumulate too much precipitation in the GrIS interior, it's possible the coastal ice margins receive too little snow, which may be more easily melted away to expose the underlying glacial ice. The top row of Figure~\ref{fig:alb} shows the summer snow cover bias relative to ERA5. There is no indication that the coastal regions have lower snow fractions, and more bare ice exposure, compared with the VR runs. In southeast Greenland, the coarse runs have anomalously high snow cover relative to the VR runs. In southwest and central-west Greenland, the snow cover is anomalously low, but arguably no more so than occurs in the VR runs. And in combining these results wih the precipitation biases (Figure~\ref{fig:bias}), there just isn't much support for a reduction in snowfall and snow cover around the coastal margins in the coarser runs.

The middle and bottom rows of Figure~\ref{fig:alb} show maps of summer albedo biases in the simulations, relative to the RACMO2.3p2 and the processed MODIS albedo dataset of \citeA{MODIS}. The model surface albedo is diagnosed from the incident and net shortwave radiative fluxes at the surface. While RACMO2.3p2 indicates that the albedo near the southwest/southeast coastal margins is lower in the lat-lon and quasi-uniform grids relative to the VR grids, the MODIS dataset does not show that the VR grids have a substantially different albedo from the coarse grids. We do observe a persistent positive albedo bias in the northwest in all grids, in both validation dataset, which is consistent with a low melt bias in that region \cite{LETAL2020JAMES}.}}

\begin{figure}[t]
\begin{center}
\begin{tabular}{cccc}
         \includegraphics[width=130mm]{temp_contours_diffALB.pdf}
\end{tabular}
\end{center}
\caption{1979-1998 northern hemisphere summer albedo bias (fraction), computed as the difference (top) from RACMO2.3p2, and (middle) the MODIS dataset of \citeA{MODIS}. The bottom row is 1979-1998 bias in summer snow cover (fraction), relative to ERA5. As all observational datasets in this plot are on high-resolution grids, they've all been mapped to the individual model grids, before taking differences.}
\label{fig:alb}
\end{figure}

@@

The $\tt Arctic-GrIS$ grid has the warmest summer lower troposphere of all grids, yet it has less melting than the quasi-uniform resolution SE grids. This suggests that grid resolution is somehow contributing to the melt biases in coarse grids, in a way that is not obvious from the large-scale dynamics. We propose a mechanism: coarse grids represent ablation zones using grid cells with mixed surface types, ice-covered and ice-free. The warmer ice-free patches may largely determine the mean state, leading to a warm bias over the ice-covered patches of the grid cell. This mechanism is supported by analysis of melt biases binned by grid-cell ice fraction, {\color{blue}{although further testing of this hypothesis is needed.}}
